---
title: "a3_task3_elena_ortiz"
author: "Elena Ortiz"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Red and Purple Sea Urchins - Data Exploration

In this document, I will explore red and purple sea urchin data from the Santa Barbara coast through data visualization and summary statistics.  

```{r}
urchins <- read.csv(here("data","urchins.csv")) %>% 
  clean_names()
```

### Red & Purple Sea Urchin Size: Exploratory Data 

Here I will compare sizes among different sea urchin populations. 

#### Geom_histogram 

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 15) +
  facet_wrap(~common_name) +
         labs(x = "Diameter (cm)", y = "Urchin count")
  
```

#### Geom_boxplot

```{r}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name", y = "Diameter (cm)")
```

### Red & purple sea urchin size: summary statistics

```{r}
urchins %>% 
  group_by(common_name) %>% 
  summarize(
    mean_size = mean(size),
    sd_size = sd(size),
    med_size = median(size),
    sample_size = n()) %>% 
  kbl() %>%
  kable_minimal()
  
```

#### Analysis 

We see that the means for the Purple and Red Urchin sizes are 4.47 and 6.78 cm respectively. The Purple Urchin population is normally distributed, with a standard deviation of 1.18, so there is smaller spread between samples in this population. On the other hand, the Red Urchin population shows a bimodal distribution with a positive skew, and larger standard deviation (1.89), or spread between samples in this population. 

### Purple Sea Urchin: Counts By Site/Year

#### Purple Urchin Subset 

```{r}
purple_urchins <- urchins %>%
  mutate(date = mdy(date)) %>%
  mutate(year_new = year(date)) %>% 
  filter(common_name == "Purple Urchin") %>%
  mutate(site_full = case_when(
    site %in% c("NAPL") ~ "Naples Reef",
    site %in% c("AQUE") ~ "Arroyo Quemado Reef",
    site %in% c("CARP") ~ "Carpinteria Reef",
    site %in% c("MOHK") ~ "Mohawk Reef",
    site %in% c("IVEE") ~ "Isla Vista Reef"))
```

#### Purple Urchin Summary Stats & Exploratory Graph

```{r}
purple_urchins %>% 
  count(site_full, year_new) %>% 
  ggplot(aes(x = year_new, y = n)) +
  geom_line(aes(color = site_full)) +
  labs(title = "Purple Urchin Populations", x = "Year", y = "Count", color = "Site Name") +
  scale_x_continuous(breaks = c(2008, 2010, 2012, 2014, 2016, 2018, 2020)) +
  theme_minimal()
  
```

### END

