---
title: "a3_task4_elena_ortiz"
author: "Elena Ortiz"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

## Kelp Removal & Sizes of Purple Urchins in Napels Reef 

Here, I compare the mean sizes of Purple Urchin populations based on kelp removal in the Naples Reef in January 2010. 

### Initial Data Wrangling 

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names() 

purple_urchins_NAPL <- urchins %>%
  mutate(date = mdy(date)) %>%
  mutate(year_new = year(date)) %>% 
  mutate(month_new = month(date)) %>%
  filter(common_name == "Purple Urchin", site == "NAPL", year_new == "2010", month_new == "1")
```

### Data Exploration 

Using histograms and QQ Plots I will do an initial visual exploration of the data. 

#### Exploratory histgrams: 

```{r}
ggplot(data = purple_urchins_NAPL, aes(x = size)) +
  geom_histogram(bins = 10) +
  facet_wrap(~treatment)
```

Both distributions appear to be normally distributed, with the "control" group depicting less spread than the "annual" group.

#### Quantile-quantile Plot 

```{r}
ggplot(data = purple_urchins_NAPL, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```

The "control" group follows a somewhat linear trend, with a wider spread in the tails. Similarly, the "annual" group is also generally linear, but shows a wider spread in the tails on either end than the "control" group. 

### Descriptive Statistics 

```{r}
kelp_summary <- purple_urchins_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean_diameter = mean(size),
            sd_diameter = sd(size),
            sample_size = n())

kelp_summary
```
### 2-Sample T-test

```{r}
annual_sample <- purple_urchins_NAPL %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- purple_urchins_NAPL %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x = annual_sample, control_sample)
```
The t-test shows that there is reasonable (25%) chance of randomly drawing sample populations with means at least this different if they were from populations with the same mean. This means it is possible that these samples were indeed drawn from populations with the same mean. In looking at the means, it is important to consider that they are fairly close with only a difference of approximately .19 between them. Additionally, the sample sizes were sufficiently large (~60), which could also impact this finding. More information is needed to determine the "significance" of this difference. 

The p-value here states that there is a 25% chance of drawing sample populations with means at least this different if they were from populations with the same mean by chance. 

### END