---
title: "a3_task3_elena_ortiz"
author: "Elena Ortiz"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

## Red and Purple Sea Urchins - Data Exploration

In this document, I will explore red and purple sea urchin data from the Santa Barbara coast through data visualization and summary statistics.  

```{r}
urchins <- read.csv(here("data","urchins.csv")) %>% 
  clean_names()
```

### Red & purple sea urchin size: exploratory data 

Here I will compare sizes among different sea urchin populations. 

#### Geom_histogram 

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 15) +
  facet_wrap(~common_name) +
         labs(x = "Diameter (cm)", y = "Urchin count")
  
```

#### Geom_boxplot

```{r}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name", y = "Diameter (cm)")
```

### Red & purple sea urchin size: summary statistics

```{r}
urchins %>% 
  group_by(common_name) %>% 
  summarize(
    mean_size = mean(size),
    sd_size = sd(size),
    med_size = median(size),
    sample_size = n()) %>% 
  kbl() %>%
  kable_minimal()
  
```

#### Analysis 




